---
title: "Introduction to GeoIndexR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to GeoIndexR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

**GeoIndexR** is a modern, modular, and fast R package designed for computing spectral and geospatial indices from multispectral raster data. Built natively on the powerful `terra` geospatial engine, it operates on `SpatRaster` objects while preserving all georeferencing, projection (CRS), resolution, and spatial extents.

### Key Capabilities

- **Unified high-level API**: Calculate indices using `geo_index()`, `geo_indices()`, or `geo_index_all()`.
- **Direct band computation**: Vectorized functions like `calc_ndvi()`, `calc_savi()`, `calc_evi()`, and `calc_index()`.
- **Intelligent band mapping**: Automatic layer matching, case-insensitivity, and preconfigured sensor presets (Sentinel-2, Landsat-8/9, Landsat-5/7).
- **Safe numerical calculation**: Zero divisions (e.g., $NIR + RED = 0$) and infinite values are converted to `NA` safely without crashing or emitting unhandled warnings.
- **Summary & Visualization**: Built-in descriptive metrics (`index_summary()`) and thematic color palettes (`plot_index()`).

---

## Quick Start

Let's load `GeoIndexR` and examine the bundled synthetic 6-band multispectral dataset (`example_multispectral`):

```{r setup}
library(GeoIndexR)
library(terra)

img <- get_example_data()
print(img)
names(img)
```

### Computing a Single Index

To compute the Normalized Difference Vegetation Index (NDVI):

```{r single_index}
ndvi <- geo_index(img, "NDVI")
print(ndvi)
```

The returned object is a single-layer `SpatRaster` whose layer name matches `"NDVI"`.

### Computing Multiple Indices

You can compute several indices in a single call. The output is a stacked multi-layer `SpatRaster`:

```{r multi_index}
indices_stack <- geo_indices(img, c("NDVI", "NDWI", "NDBI", "NDMI"))
print(indices_stack)
names(indices_stack)
```

### Automatic Computation of All Eligible Indices

If you have an arbitrary multispectral image and want to discover and compute all indices compatible with the available bands:

```{r all_indices}
all_indices <- geo_index_all(img, quiet = FALSE)
names(all_indices)
```

---

## Exploring the Index Registry

All supported indices are documented in a centralized registry accessible with `index_registry()`:

```{r registry}
reg <- index_registry()
reg[, c("index", "name", "category", "required_bands")]
```

You can filter by category (e.g., `"water"`, `"vegetation"`, `"urban"`):

```{r water_registry}
index_registry(category = "water")[, c("index", "name", "formula")]
```

---

## Summary Statistics

The `index_summary()` function computes cell-level summary statistics including minimum, maximum, mean, median, standard deviation, and missing data percentage:

```{r summary}
index_summary(indices_stack)
```
